Gammapy: A Python package for gamma-ray astronomy

### 2023-06-01 src/ms-review.tex
Add CTA prod5 reference
Move authors to direct contribution to paper

\author{
	{\it Paper Authors} \and
    Axel Donath \inst{\ref{inst:28}} \and
	Régis Terrier \inst{\ref{inst:20}} \and
	Quentin Remy \inst{\ref{inst:17}} \and
	Atreyee Sinha \inst{\ref{inst:15}} \and
	Cosimo Nigro \inst{\ref{inst:24}} \and
	Fabio Pintore \inst{\ref{inst:26}} \and
	Bruno Khélifi \inst{\ref{inst:20}} \and
	Laura Olivera-Nieto \inst{\ref{inst:17}}\and
	Jose Enrique {Ruiz} \inst{\ref{inst:23}} \and
	\\
	{\it Coordination Committee}\thanks{Corresponding author:
\href{mailto:GAMMAPY-COORDINATION-L@IN2P3.FR}{GAMMAPY-COORDINATION-L@IN2P3.FR}} \and
	Fabio Acero \inst{\ref{inst:1}} \and
	David Berge \inst{\ref{inst:6}, \ref{inst:7}} \and
	Catherine Boisson \inst{\ref{inst:9}} \and
	Jose Louis {Contreras} \inst{\ref{inst:15}} \and
	Axel Donath \inst{\ref{inst:28}} \and
	Stefan Funk \inst{\ref{inst:19}} \and
	Christopher {van Eldik} \inst{\ref{inst:19}} \and
	Matthias Fü{\ss}ling \inst{\ref{inst:18}} \and
	Jim Hinton \inst{\ref{inst:17}} \and
	Bruno Khélifi \inst{\ref{inst:20}} \and
	Rub{\'e}n L{\'o}pez-Coto \inst{\ref{inst:23}} \and
	Fabio Pintore \inst{\ref{inst:26}} \and
	Régis Terrier \inst{\ref{inst:20}} \and
	Roberta Zanin \inst{\ref{inst:18}} \and
    \\
	{\it Gammapy Project Contributors} \and
	Arnau Aguasca-Cabot \inst{\ref{inst:2}, \ref{inst:3}, \ref{inst:4}, \ref{inst:5}} \and
	Pooja Bhattacharjee \inst{\ref{inst:8}} \and
	Kai Brügge \inst{\ref{inst:10},\ref{inst:11}} \and
	Johannes Buchner \inst{\ref{inst:12}} \and
	David {Carreto Fidalgo} \inst{\ref{inst:13}} \and
	Andrew Chen \inst{\ref{inst:14}} \and
	Mathieu {de Bony de Lavergne} \inst{\ref{inst:8}} \and
	José Vinícius {de Miranda Cardoso}\inst{\ref{inst:16}} \and
	Christoph Deil \inst{\ref{inst:17}} \and
	Luca Giunti \inst{\ref{inst:20}} \and
	Léa Jouvin \inst{\ref{inst:32}} \and
	Johannes King \inst{\ref{inst:30}, \ref{inst:17}} \and
	Julien Lefaucheur \inst{\ref{inst:21}, \ref{inst:20}} \and
	Marianne Lemoine-Goumard \inst{\ref{inst:31}} \and
	Jean-Philippe Lenain \inst{\ref{inst:22}} \and
	Maximilian Linhoff \inst{\ref{inst:11}} \and
	Lars Mohrmann \inst{\ref{inst:17}} \and
	Daniel Morcuende \inst{\ref{inst:15}} \and
	Sebastian Panny \inst{\ref{inst:25}} \and
	Maxime Regeard \inst{\ref{inst:20}} \and
	Lab Saha \inst{\ref{inst:15}} \and
	Hubert Siejkowski \inst{\ref{inst:27}} \and
	Aneta Siemiginowska \inst{\ref{inst:28}} \and
	Brigitta M. {Sipőcz} \inst{\ref{inst:29}} \and
	Tim Unbehaun \inst{\ref{inst:19}} \and
	Thomas Vuillaume \inst{\ref{inst:8}}
}

@@

In addition to the instrument characteristics described above, there is also the instrumental
background, that results from hadronic events being misclassified as \gammaray events. These
events constitute a uniform background to the \gammaray events For \fermi the residual hadronic background
is very small ($<1\%$) because of its veto layer and can often be neglected. For IACTs and WCTs in contrast
the background can account for a very large part ($>95\%$) of the events and must be treated
accordingly in the analysis. As the background is very specific to the instrument, \gammapy typically 
relies on the background models provided with the DL3 data. The background typically only depends
on the reconstructed event positon and energy $Bkg(p, E)$. However in general all IRFs depend on the
geometrical parameters of the detector, e.g. location of an event in the FoV or the elevation
angle of the incoming direction of the event. Consequently IRFs might be parametrised as functions
of detector specific coordinates as well.

@@

 \subsubsection{DL4: Binned Data Structures}
 The counts data and the reduced IRFs in the form of maps are bundled into datasets
 that represent the fourth data level (DL4). These reduced datasets can be written to disk,
 in a format specific to \gammapy to allow users to read them back at any time later
 for modelling and fitting. Different variations of such datasets support different 
 analysis methods and fit statistics. The datasets can be used to perform a joint-likelihood
 fit, allowing one to combine different measurements, e.g. from different observations
 but also from different instruments or event classes. They can also be used for binned
 simulation as well as event sampling to simulate DL3 events data.

@@

\subsection{From DL4 to DL5/6: Modeling and Fitting}
The next step is then typically to model the datasets using binned Poisson maximum likelihood fitting.
Assuming Poisson statistics per bin the log-likelihood of an observation is given by the Cash statistics\citep{Cash}:
%
\begin{equation}
	\mathcal{C} = \sum_{i=0}^K N_{\mathrm{Pred, i}} - N_{\mathrm{Obs, i}} \log{N_{\mathrm{Pred, i}}} 
\end{equation}

Where the expected number of events in a \gammaray observation is given by:
 \begin{equation}
	 \begin{split}
		N_{\mathrm{Pred}}(p, E; \hat{\theta})\ {\rm d}p\ {\rm d}E =  &E_{\rm disp} \circledast \left[ PSF \circledast \left( A_{\rm eff} \cdot t_{\rm obs} \cdot \Phi(\hat{\theta}) \right) \right]\\
						& + Bkg(p, E) \cdot t_{\rm obs}
	 \end{split}
 \end{equation}

With the IRF components described in the previous section as well as an analytical model to describe the
intensity of the radiation from gamma-ray sources as a function of the energy, $E_{\rm true}$,
and of the position in the FoV, $p_{\rm true}$:
%
\begin{equation}
    \Phi(p_{\rm true}, E_{\rm true}; \hat{\theta}), [\Phi] = \si{TeV^{-1}.cm^{-2}.s^{-1}}
    \label{eq:general_source_model}
\end{equation}
Where $\hat{\theta}$ is a set of model parameters that can be adjusted in a fit. 
Observtaions can be either modelled individually, or in a joint likelihood analysis.
In the latter case the total (jpint) log-likleihood is given by the sum of the log-likelihoods
per observation $\mathcal{L} = \sum{i=0}^M \mathcal{C_i}$ for M observations. Other fit
statistics such as \textit{WStat} or a simple $\chi^2$ for flux points are also often used
and are supported by \gammapy as well. Models can be simple analytical models or more complex
ones from radiation mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{0}$ decay).

@@

Parametric models and all the functionality realted to fitting is implemented in 
\code{gammapy.modeling} and \code{gammapy.estimators}, where the latter is used to
compute higher level science products such as flux and significance maps,
light curves or flux points.

>>>>

\author{
	{\it Paper Authors} \and
    Axel Donath \inst{\ref{inst:28}} \and
	Régis Terrier \inst{\ref{inst:20}} \and
	Quentin Remy \inst{\ref{inst:17}} \and
	Atreyee Sinha \inst{\ref{inst:15}} \and
	Cosimo Nigro \inst{\ref{inst:24}} \and
	Fabio Pintore \inst{\ref{inst:26}} \and
	Bruno Khélifi \inst{\ref{inst:20}} \and
	Laura Olivera-Nieto \inst{\ref{inst:17}}\and
	Jose Enrique {Ruiz} \inst{\ref{inst:23}} \and
	Kai Brügge \inst{\ref{inst:10},\ref{inst:11}} \and
	Maximilian Linhoff \inst{\ref{inst:11}} \and
	Jose Louis {Contreras} \inst{\ref{inst:15}} \and
	\\
	{\it Coordination Committee}\thanks{Corresponding author:
\href{mailto:GAMMAPY-COORDINATION-L@IN2P3.FR}{GAMMAPY-COORDINATION-L@IN2P3.FR}} \and
	Fabio Acero \inst{\ref{inst:1}} \and
	David Berge \inst{\ref{inst:6}, \ref{inst:7}} \and
	Catherine Boisson \inst{\ref{inst:9}} \and
	Jose Louis {Contreras} \inst{\ref{inst:15}} \and
	Axel Donath \inst{\ref{inst:28}} \and
	Stefan Funk \inst{\ref{inst:19}} \and
	Christopher {van Eldik} \inst{\ref{inst:19}} \and
	Matthias Fü{\ss}ling \inst{\ref{inst:18}} \and
	Jim Hinton \inst{\ref{inst:17}} \and
	Bruno Khélifi \inst{\ref{inst:20}} \and
	Rub{\'e}n L{\'o}pez-Coto \inst{\ref{inst:23}} \and
	Fabio Pintore \inst{\ref{inst:26}} \and
	Régis Terrier \inst{\ref{inst:20}} \and
	Roberta Zanin \inst{\ref{inst:18}} \and
    \\
	{\it Gammapy Project Contributors} \and
	Arnau Aguasca-Cabot \inst{\ref{inst:2}, \ref{inst:3}, \ref{inst:4}, \ref{inst:5}} \and
	Pooja Bhattacharjee \inst{\ref{inst:8}} \and
	Johannes Buchner \inst{\ref{inst:12}} \and
	David {Carreto Fidalgo} \inst{\ref{inst:13}} \and
	Andrew Chen \inst{\ref{inst:14}} \and
	Mathieu {de Bony de Lavergne} \inst{\ref{inst:8}} \and
	José Vinícius {de Miranda Cardoso}\inst{\ref{inst:16}} \and
	Christoph Deil \inst{\ref{inst:17}} \and
	Luca Giunti \inst{\ref{inst:20}} \and
	Léa Jouvin \inst{\ref{inst:32}} \and
	Johannes King \inst{\ref{inst:30}, \ref{inst:17}} \and
	Julien Lefaucheur \inst{\ref{inst:21}, \ref{inst:20}} \and
	Marianne Lemoine-Goumard \inst{\ref{inst:31}} \and
	Jean-Philippe Lenain \inst{\ref{inst:22}} \and
	Lars Mohrmann \inst{\ref{inst:17}} \and
	Daniel Morcuende \inst{\ref{inst:15}} \and
	Sebastian Panny \inst{\ref{inst:25}} \and
	Maxime Regeard \inst{\ref{inst:20}} \and
	Lab Saha \inst{\ref{inst:15}} \and
	Hubert Siejkowski \inst{\ref{inst:27}} \and
	Aneta Siemiginowska \inst{\ref{inst:28}} \and
	Brigitta M. {Sipőcz} \inst{\ref{inst:29}} \and
	Tim Unbehaun \inst{\ref{inst:19}} \and
	Thomas Vuillaume \inst{\ref{inst:8}}
}

@@

In addition to the instrument characteristics described above, there is also the instrumental
background, that results from hadronic events being misclassified as \gammaray events. These
events constitute a uniform background to the \gammaray events. For \fermi the residual hadronic background
is very small ($<1\%$) because of its veto layer and can often be neglected. For IACTs and WCDs in contrast,
the background can account for a very large part ($>95\%$) of the events and must be treated
accordingly in the analysis. As the background is very specific to the instrument, \gammapy typically 
relies on the background models provided with the DL3 data. The background typically only depends
on the reconstructed event position and energy $Bkg(p, E)$. However in general all IRFs depend on the
geometrical parameters of the detector, e.g. location of an event in the FoV or the elevation
angle of the incoming direction of the event. Consequently IRFs might be parametrised as functions
of detector specific coordinates as well.

@@

 \subsection{DL4: Binned Data Structures}
 The counts data and the reduced IRFs in the form of maps are bundled into datasets
 that represent the fourth data level (DL4). These reduced datasets can be written to disk,
 in a format specific to \gammapy to allow users to read them back at any time later
 for modelling and fitting. Different variations of such datasets support different 
 analysis methods and fit statistics. The datasets can be used to perform a joint-likelihood
 fit, allowing one to combine different measurements, e.g. from different observations
 but also from different instruments or event classes. They can also be used for binned
 simulation as well as event sampling to simulate DL3 events data.

@@

\subsection{From DL4 to DL5/6: Modeling and Fitting}
The next step is then typically to model the datasets using binned Poisson maximum likelihood fitting.
Assuming Poisson statistics per bin the log-likelihood of an observation is given by the Cash statistics\citep{Cash}:
%
\begin{equation}
	\mathcal{C} = 2 \sum_{i=0}^K N_{\mathrm{Pred, i}} - N_{\mathrm{Obs, i}} \log{N_{\mathrm{Pred, i}}} 
\end{equation}

Where the expected number of events for the observation is given by forward folding a source model through
the instrument response:
 \begin{equation}
	 \begin{split}
		N_{\mathrm{Pred}}(p, E; \hat{\theta})\ {\rm d}p\ {\rm d}E =  &E_{\rm disp} \circledast \left[ PSF \circledast \left( A_{\rm eff} \cdot t_{\rm obs} \cdot \Phi(\hat{\theta}) \right) \right]\\
						& + Bkg(p, E) \cdot t_{\rm obs}
	 \end{split}
 \end{equation}

The equation includes the IRF components described in the previous section, as well as an analytical model to describe the
intensity of the radiation from \gammaray sources as a function of the energy, $E_{\rm true}$,
and of the position in the FoV, $p_{\rm true}$:
%
\begin{equation}
    \Phi(p_{\rm true}, E_{\rm true}; \hat{\theta}), [\Phi] = \si{TeV^{-1}.cm^{-2}.s^{-1}}
    \label{eq:general_source_model}
\end{equation}

Where $\hat{\theta}$ is a set of model parameters that can be adjusted in a fit. 
Observations can be either modeled individually, or in a joint likelihood analysis.
In the latter case the total (joint) log-likelihood is given by the sum of the log-likelihoods
per observation $\mathcal{L} = \sum{i=0}^M \mathcal{C_i}$ for M observations. Other fit
statistics such as \textit{WStat}, where the background is estimated from an independent
measurement or a simple $\chi^2$ for flux points are also often used
and are supported by \gammapy as well. Models can be simple analytical models or more complex
ones from radiation mechanisms of accelerated particle populations (e.g. inverse Compton or $\pi^{0}$ decay).

@@

Parametric models and all the functionality related to fitting is implemented in 
\code{gammapy.modeling} and \code{gammapy.estimators}, where the latter is used to
compute higher level science products such as flux and significance maps,
light curves or flux points.
